---
title: "Homework 4"
author: "Dakota Wilson"
date: "2/20/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ISLR2)
weekly = ISLR2::Weekly
```

### Question 4

**a) Suppose that we have a set of observations, each with measurements on p = 1 feature, X. We assume that X is uniformly (evenly) distributed on [0, 1]. Associated with each observation is a response value. Suppose that we wish to predict a test observation’s response using only observations that are within 10 % of the range of X closest to that test observation. For instance, in order to predict the response for a test observation with X = 0.6, we will use observations in the range [0.55,0.65]. On average, what fraction of the available observations will we use to make the prediction?**

We would expect slightly less than 10% of the avaliable observations used in each prediciton. This would be exactly 10%, except for on the edges (<.05 and >.95), where there would not be enough room one one side to collect all of the needed 5%. This comes out to exactly:

```{r}
(.90*.1)+(.02*.09)+(.02*.08)+(.02*.07)+(.02*.06)+(.02*.05)
```
9.7% of the time

**b) Now suppose that we have a set of observations, each with measurements on p = 2 features, X1 and X2. We assume that (X1, X2) are uniformly distributed on [0, 1] × [0, 1]. We wish to predict a test observation’s response using only observations that are within 10 % of the range of X1 and within 10 % of the range of X2 closest to that test observation. For instance, in order to predict the response for a test observation with X1 = 0.6 and X2 = 0.35, we will use observations in the range [0.55, 0.65] for X1 and in the range [0.3,0.4] for X2. On average, what fraction of the available observations will we use to make the prediction?**

If there are now two directions, we can multiply the result of the first portion by itself. This would simply be .097*.097 = .009409 which is rougly equal to 1%.

**c) Now suppose that we have a set of observations on p = 100 features. Again the observations are uniformly distributed on each feature, and again each feature ranges in value from 0 to 1. We wish to predict a test observation’s response using observations within the 10 % of each feature’s range that is closest to that test observation. What fraction of the available observations will we use to make the prediction?**


This is the same principal as before, except now it is to the power of 100. This comes out to 4.755251e-102, which is so small it can be assumed the percentage is zero. This means no observations will be available for prediction.

**d) Using your answers to parts (a)–(c), argue that a drawback of KNN when p is large is that there are very few training observations “near” any given test observation.**

When p is large, there is a low amount of available observations for prediction, and the availability continues to decrease the larger that p is.

**e) Now suppose that we wish to make a prediction for a test observation by creating a p-dimensional hypercube centered around the test observation that contains, on average, 10 % of the training observations. For p = 1,2, and 100, what is the length of each side of the hypercube? Comment on your answer.**

This would be set up similarly to the last four parts, except instead of using 9.7%, we would just use 10%. This gives us 10% for p = 1, 1% for p = 2 and 1e-98% for p = 100.

### Question 13

**This question should be answered using the Weekly data set, which is part of the ISLR2 package. This data is similar in nature to the Smarket data from this chapter’s lab, except that it contains 1, 089 weekly returns for 21 years, from the beginning of 1990 to the end of 2010.**


**(a) Produce some numerical and graphical summaries of the Weekly data. Do there appear to be any patterns?**

```{r}
summary(weekly)
```

**(b) Use the full data set to perform a logistic regression with Direction as the response and the five lag variables plus Volume as predictors. Use the summary function to print the results. Do any of the predictors appear to be statistically significant? If so, which ones?**

```{r}
glm.fit = glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume, data = weekly, family = binomial)
summary(glm.fit)
```

Lag2 appears to be statistically significant at the .01 level.

**(c) Compute the confusion matrix and overall fraction of correct predictions. Explain what the confusion matrix is telling you about the types of mistakes made by logistic regression.**

```{r}
glm.probs <- predict(glm.fit, type = "response")
contrasts(weekly$Direction)
glm.pred <- rep("Down", length(weekly$Direction))
glm.pred[glm.probs > .5] = "Up"
table(glm.pred, weekly$Direction)
(557+54)/1089
mean(glm.pred ==  weekly$Direction)
```
This means that only 56.1% of my predictions are considered accurate.

**(d) Now fit the logistic regression model using a training data period from 1990 to 2008, with Lag2 as the only predictor. Compute the confusion matrix and the overall fraction of correct predictions for the held out data (that is, the data from 2009 and 2010).**

```{r}
glm.fit = glm(Direction ~ Lag2, data = weekly, family = binomial)
summary(glm.fit)
glm.probs <- predict(glm.fit, type = "response")
contrasts(weekly$Direction)
glm.pred <- rep("Down", length(weekly$Direction))
glm.pred[glm.probs > .5] = "Up"
table(glm.pred, weekly$Direction)
(579+33)/1089
mean(glm.pred ==  weekly$Direction)
```

**Repeat (d) using LDA**

```{r}

```

**(f) Repeat (d) using QDA.**



**(g) Repeat (d) using KNN with K = 1.**



**(h) Repeat (d) using naive Bayes.**



**(i) Which of these methods appears to provide the best results on this data?**



**(j) Experiment with different combinations of predictors, including possible transformations and interactions, for each of the methods. Report the variables, method, and associated confusion matrix that appears to provide the best results on the held out data. Note that you should also experiment with values for K in the KNN classifier.**